---
title: "hw02-Robby-Grewal Basics of Data Frames"
output: github_document
---


# Import the Data in R_
```{r}
library(dplyr)
```

#### .csv

```{r}

dat1 <- read.csv('~/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv',
                 colClasses = c(Player="character",
                               Team="character",
                               Experience="character",
                               Position= "factor", 
                               Salary="double"))

# Check the Structure
str(dat1)
```

#### _csv

```{r}
library(readr)
dat2 <- read_csv('~/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv',
                 col_types = list(Player = col_character(),
                                  Team = col_character(), 
                                  Experience = col_character(), 
                                  Position = col_factor(c(
                                    "C", "PF", "PG","SF", "SG")), 
                                  Salary= col_double()))

str(dat2)

```

# Right After Importing the Data_

#### Replacing R with 0 and Integer Conversion in .csv and _csv

```{r}
dat1$Experience <- as.integer(replace(dat1$Experience,
                                     dat1$Experience == 'R', 0))
```

```{r}
dat2$Experience <-  as.integer(replace(dat2$Experience,
                                       dat2$Experience == 'R', 0))
```

# Performance of Players_

##### Adding Variables and EFF

```{r}
dat1$Missed_FG <- dat1$FGA - dat1$FGM
dat1$Missed_FT <- dat1$FTA - dat1$FTM
dat1$PTS <- (dat1$Points3 *3 ) + (dat1$Points2 * 2) + dat1$FTM
dat1$REB <- dat1$OREB + dat1$DREB
dat1$MPG <- (dat1$MIN / dat1$GP)
dat1$EFF <- ((dat1$PTS + dat1$REB + dat1$AST + 
              dat1$STL+ dat1$BLK - dat1$Missed_FG -
             dat1$Missed_FT - dat1$TO) / dat1$GP)


```


#### Summary of EFF

```{r}
summary(dat1$EFF)
```


#Histogram of EFF & Rest of This Problem

```{r}
hist(dat1$EFF, xlab = "EFF", ylab = "Frequency", 
     main = "Histogram of Efficiency", col= "blue")

# Top 10 by decreasing EFF
top_ten <- dat1[order(-dat1$EFF),] [1:10,c("Player", "Team", "Salary","EFF")]
top_ten

# Negative EFF players
filter(dat1, EFF < 0)

# Use the function cor() to compute the correlation coefficients between EFF and all the variables used in the EFF formula
PTS = cor(dat1$EFF, dat1$PTS)
REB = cor(dat1$EFF, dat1$REB)
AST = cor(dat1$EFF, dat1$AST)
STL = cor(dat1$EFF, dat1$STL)
BLK = cor(dat1$EFF, dat1$BLK)
Missed_FG = cor(dat1$EFF, -dat1$Missed_FG)
Missed_FT = cor(dat1$EFF, -dat1$Missed_FT)
TO = cor(dat1$EFF, dat1$TO)
GP = cor(dat1$EFF, dat1$GP)

# Correlation Between Player Stats and EFF
correlation = c(PTS = cor(dat1$EFF, dat1$PTS), 
                REB = cor(dat1$EFF, dat1$REB), 
                AST = cor(dat1$EFF, dat1$AST),
                STL = cor(dat1$EFF, dat1$STL),
                BLK = cor(dat1$EFF, dat1$BLK),
                Missed_FG = -cor(dat1$EFF, dat1$Missed_FG),
                Missed_FT = -cor(dat1$EFF, dat1$Missed_FT),
                TO = -cor(dat1$EFF, dat1$TO),
                GP = cor(dat1$EFF, dat1$GP))

sorted_correlation = sort(correlation, decreasing = TRUE)
barplot(sorted_correlation, main = 'Correlation between Player Stats and EFF', cex.names = .60)

```

# Efficiency and Salary
```{r}
#Scatterplot
plot(dat1$EFF, dat1$Salary, main = "Efficiency and Salary", xlab = "Efficiency", ylab =  "Salary (in Millions)")
lines(lowess(dat1$EFF, dat1$Salary), col = 'red')

#Correlation
cor(dat1$EFF, dat1$Salary)
#There is a positive correlation between player efficiency and salary

#Player Data Frame
players2 <- dat1[dat1$MPG > 20,]


#Scatterplot
plot(players2$EFF, players2$Salary, main = "Efficiency and Salary in More Established Players", xlab = "Efficiency", ylab = "Salary (in Millions)")
lines(lowess(players2$EFF, players2$Salary), col = 'red')
cor(players2$EFF, players2$Salary)
#Less correlation between salary and efficiency of more established players
```


#Comments and Reflection_
Overall, this homework was pretty straightforward, but it was a little time consuming for me to look up all the correct ways to use the functions, as I still don't feel like I have a very natural grasp of R at this point. It took me almost 8 hours to complete, and I worked with friends in the class to figure out how to tackle certain aspects such as the read.csv and read_csv functions. 











